---
title: Deskriptivní statistiky časových řad
author: Daniel Bartušek
output: html_document

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---
```{r include = F}
rm(list = ls())
options(scipen = 100, digits = 8)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(readr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(dse)
library(readxl)
library(zoo)
library(tseries)
library(xts)
library(imputeTS)
library(ggpubr)
library(corrplot)
library(kableExtra)
library(stargazer)
```

```{r}
# lag function
lagpad <- function(x, k = 1) {
  i <- is.vector(x)
  if (is.vector(x)) x <- matrix(x) else x <- matrix(x, nrow(x))
  if (k > 0) {
    x <- rbind(matrix(rep(NA, k * ncol(x)), ncol = ncol(x)), matrix(x[1:(nrow(x) - k), ], ncol = ncol(x)))
  } else {
    x <- rbind(matrix(x[(-k + 1):(nrow(x)), ], ncol = ncol(x)), matrix(rep(NA, -k * ncol(x)), ncol = ncol(x)))
  }
  if (i) x[1:length(x)] else x
}
```

## Load data

```{r}
nat_gas_spot <- read_excel("data/NG_PRI_FUT_S1_D.xls", sheet = "Data 1", skip = 2) %>%
  rename("Natural_Gas" = 2)
crude_oil_spot <- read_excel("data/PET_PRI_SPT_S1_D.xls", sheet = "Data 1", skip = 2) %>%
  rename("Crude_Oil_WTI" = 2, "Crude_Oil_Europe_Brent" = 3)
gasoline <- read_excel("data/PET_PRI_SPT_S1_D.xls", sheet = "Data 2", skip = 2) %>%
  rename("NY_Gasoline" = 2, "Gulf_Gasoline" = 3)
heating_oil <- read_excel("data/PET_PRI_SPT_S1_D.xls", sheet = "Data 4", skip = 2) %>%
  rename("Heating_Oil" = 2)
diesel <- read_excel("data/PET_PRI_SPT_S1_D.xls", sheet = "Data 5", skip = 2) %>%
  select(-2, -3) %>%
  rename("Diesel" = 2)
```

### Group to dataframe and time-series form

```{r include=F}
dt <- nat_gas_spot %>%
  full_join(crude_oil_spot, by = "Date") %>%
  full_join(gasoline, by = "Date") %>%
  full_join(heating_oil, by = "Date") %>%
  full_join(diesel, by = "Date") %>%
  arrange(Date) %>%
  filter(Date >= "1997-01-07") %>%
  mutate(Date=as.Date(Date))

nat_gas_spot_ts <- xts(nat_gas_spot$Natural_Gas, frequency = 365, order.by = as.Date(nat_gas_spot$Date))
crude_oil_spot <- ts(crude_oil_spot[, 1:2], frequency = 7, start = as.Date("1997-01-07"), end = as.Date("2021-07-01"))
NY_Gasoline_ts <- ts(gasoline[, c(1, 2)], start = as.Date("1997-01-07"))
Gulf_Gasoline_ts <- ts(gasoline[, c(1, 3)], start = as.Date("1997-01-07"))
heating_oil_ts <- ts(heating_oil, start = as.Date("1997-01-07"))
diesel_ts <- ts(diesel, start = as.Date("1997-01-07"))
ts_list <- list(nat_gas_spot_ts, crude_oil_spot, NY_Gasoline_ts, Gulf_Gasoline_ts, heating_oil_ts, diesel_ts)
```

```{r}
attr(nat_gas_spot_ts, "frequency") <- 7
nat_gas_spot_ts <- na.omit(nat_gas_spot_ts)
plot(decompose(as.ts(nat_gas_spot_ts))) 
```
Dont know hot to use this decomposition, but might be a useful tool in the future.


### Removing NAs

```{r}
colSums(is.na(dt))
nrow(dt) # about 2% NAs
aux <- dt[!complete.cases(dt), ]
head(aux) %>% 
  kbl() %>%
  kable_classic_2(full_width = F)
```

We see that the Europe Brent crude oil benchmark is reported on days when the others arent, so we only use WTI.
Then we will delete rows filled with NAs.

```{r}
dt <- dt %>% select(-Crude_Oil_Europe_Brent)
NA_dates <- dt %>%
  filter_at(vars(-Date), all_vars(is.na(.))) %>%
  pull(Date)
dt <- dt %>% filter(!Date %in% NA_dates)

aux <- dt[!complete.cases(dt), ]
aux[10:20, ]  %>% 
  kbl() %>%
  kable_classic_2(full_width = F)

#Number of NAs in each column:
colSums(is.na(aux))
dt <- dt %>% mutate_all(~ replace(., is.na(.), mean(., na.rm = TRUE)))
```
```{r echo =F}
print(paste("Number of rows in the dataframe:",nrow(dt)))
```

Data still not complete, mostly differs in the data sources (1st column vs rest). It might be best to treat the first and second source as separate, but there is only a few NAs, so for this purpose, we just substitute with mean. Could also substitue with rolling average.

### Summary statistics - Prices

```{r}
long_dt <- melt(dt, id.vars = c("Date"))

ggplot(long_dt, aes(factor(variable), value)) + 
geom_boxplot() + 
facet_wrap(~variable, scale="free")

long_dt %>% group_by(as.factor(variable)) %>%
  get_summary_stats(type = "common")%>%
  kbl() %>%
  kable_classic_2(full_width = F)

```

```{r}
cor_dt <- cor(dt[,-1])
corrplot(cor_dt, type="upper", order="hclust", tl.col="black", tl.srt=45)
```

We see that this **crude oil WTI metric** had negative prices at one point. This would not let us do log-returns, so we substitute with low value for now. Should be treated more rigorously later on.

```{r}
dt[dt$Crude_Oil_WTI<0,"Crude_Oil_WTI"]
dt[dt$Date=="2020-04-20","Crude_Oil_WTI"]<-0.01
```

### Log returns

```{r}
log_ret_dt <- tibble(dt[2:nrow(dt), "Date"])

for (i in 2:(ncol(dt))) {
  var_name <- paste(colnames(dt)[i], "log_ret", sep = "_")
  print(var_name)
  vals <- log(dt[, i]) %>% pull()
  lag_vals <- diff(vals, lag = 1)
  log_ret_dt[var_name] <- lag_vals
}

```

### Summary statistics - Log Returns

```{r}
long_log_ret <- melt(log_ret_dt, id.vars = c("Date"))
head(long_dt)%>%
    kbl() %>%
  kable_classic_2(full_width = F)


ggplot(long_log_ret, aes(factor(variable), value)) + 
geom_boxplot() + 
facet_wrap(~variable, scale="free")

long_log_ret %>% group_by(as.factor(variable)) %>%
  get_summary_stats(type = "common")%>%
    kbl() %>%
  kable_classic_2(full_width = F)

```

```{r}
cor_long_dt <- cor(log_ret_dt[,-1])
corrplot(cor_long_dt, type="upper", order="hclust", tl.col="black", tl.srt=45)
```

### Basic plots

```{r}
ggplot(long_dt, aes(x = Date, y = value)) +
  geom_line() +
  facet_wrap(~variable, scales = "free") +
  theme(axis.text.x = element_blank()) +
  ggtitle("Prices") +
  geom_vline(xintercept = as.POSIXct("2008-9-15"), color = "Red")
```

Everything but Natural gas looks very correlated, price data are obviously not stationary, so we transform to log-returns.

```{r}
ggplot(long_log_ret, aes(x = Date, y = value)) +
  geom_line() +
  facet_wrap(~variable, scales = "free") +
  theme(axis.text.x = element_blank()) +
  ggtitle("Log Returns - May be stationary")
```

## Statistical Tests

### Autocorrelation tests

```{r natgas,echo = F}
#Natural Gas
model <- lm(Natural_Gas_log_ret ~ I(lagpad(Natural_Gas_log_ret, 1)), data = log_ret_dt)
res <- summary(model)$residuals
ar_lm_nat_gas <- lm(res ~ lagpad(res, 1))
```
```{r adf_natgas}
adf.test(log_ret_dt$Natural_Gas_log_ret) #no unit root
```

```{r natgas_res,echo=TRUE,results='asis'}
stargazer(ar_lm_nat_gas, header=FALSE,type="html") #no autocorrelation of residuals
```

```{r crd,echo = F}
#Crude Oil
model <- lm(Crude_Oil_WTI_log_ret ~ I(lagpad(Crude_Oil_WTI_log_ret, 1)), data = log_ret_dt)
res <- summary(model)$residuals
ar_lm_crd_oil<- lm(res ~ lagpad(res, 1))
```

```{r crd_adf}
adf.test(log_ret_dt$Crude_Oil_WTI_log_ret) # no unit root
```

```{r crd_res,echo=TRUE,results='asis'}
stargazer(ar_lm_crd_oil, header=FALSE,type="html")  #AR(1) OF RESIDUALS
```


```{r NY_gas,echo = F}
#NY Gasoline
model <- lm(NY_Gasoline_log_ret ~ I(lagpad(NY_Gasoline_log_ret, 1)), data = log_ret_dt)
res <- summary(model)$residuals
ar_lm_gasoline<- lm(res ~ lagpad(res, 1))
```

```{r NY_gas_adf}
adf.test(log_ret_dt$NY_Gasoline_log_ret) # no unit root
```

```{r NY_gas_res,echo=TRUE,results='asis'}
stargazer(ar_lm_gasoline, header=FALSE,type="html")  #AR(1) OF RESIDUALS
```

```{r gulf_gas,echo = F}
model <- lm(Gulf_Gasoline_log_ret ~ I(lagpad(Gulf_Gasoline_log_ret, 1)), data = log_ret_dt)
res <- summary(model)$residuals
ar_lm_gulf_gas<- lm(res ~ lagpad(res, 1))
```

```{r gulf_gas_adf}
adf.test(log_ret_dt$Gulf_Gasoline_log_ret) # no unit root
```

```{r gulf_gas_res,echo=TRUE,results='asis'}
stargazer(ar_lm_gulf_gas, header=FALSE,type="html") #AR(1) OF RESIDUALS
```

```{r heating_oil,echo = F}
model <- lm(Heating_Oil_log_ret ~ I(lagpad(Heating_Oil_log_ret, 1)), data = log_ret_dt)
res <- summary(model)$residuals
ar_lm_heating_oil<- lm(res ~ lagpad(res, 1))
```

```{r heating_oil_adf}
adf.test(log_ret_dt$Heating_Oil_log_ret) # no unit root
```

```{r heating_oil_res,echo=TRUE,results='asis'}
stargazer(ar_lm_heating_oil, header=FALSE,type="html") #AR(1) OF RESIDUALS
```

```{r diesel,echo = F}
model <- lm(Diesel_log_ret ~ I(lagpad(Diesel_log_ret, 1)), data = log_ret_dt)
res <- summary(model)$residuals
ar_lm_diesel<- lm(res ~ lagpad(res, 1))
```

```{r diesel_adf}
adf.test(log_ret_dt$Diesel_log_ret) # no unit root
```

```{r diesel_res,echo=TRUE,results='asis'}
stargazer(ar_lm_diesel, header=FALSE,type="html")  #no autocorrelation of residuals
```

Few of these commodities have autocorrelation of residuals, which should be treated by FGLS or using robust standard errors.

## MA(1) Test
Only for Diesel since the others would be similar.
MA is obviously good fit since the data are already differenced (endogenously created)

```{r ma,echo = TRUE,results='asis'}
diesel_ma <- arima(log_ret_dt$Diesel_log_ret, order = c(0, 0, 1))
stargazer(diesel_ma, header=FALSE,type="html")
ts.plot(log_ret_dt$Diesel_log_ret)
points(log_ret_dt$Diesel_log_ret-diesel_ma$residuals, type = "l", col = 2, lty = 2)
```


```{r ar_differently,echo=TRUE,results='asis'}
ar<-arima(log_ret_dt$Diesel_log_ret, order = c(1, 0, 0))
stargazer(ar, header=FALSE,type="html")
```
Same as summary(model) from before, so I could also use this package for AR testing.
